---
title: "Initial Korean covid 19"
author: "Karla Cota Lopez"
date: "October 10, 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
 
# Your Question - Is dying independent of sex?

#Methods
Result: Categorical
Sex~Categorical

We use the Cat~Cat helper file


$H_0$: Dying is not independent of sex
$H_A$: Dying is independent of sex


## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  
The bar chart shows that men die at a higher rate compared to women.




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
Among the people who died 64% were men and 36% were female. The men percentage is double the female percentage.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi-squared test was 0.0012. This is less than 0.05, thus we reject the null hypothesis. Men are 2.3 times more likely to die of covid in korea than the women are. 


## Conclusions
Our graphical results showed us that men are more likely to pass from covid in korea than women are. Among the people who died in Jan-May 2020, 64% where men and 36% were women. The p-value for the Chi-squared test was less than .05, making us reject the null hypothesis. We conclude that there is a difference in dying based on sex. Our data showed that men are 2.3 times more likely to die from covid in korea than women.  
